---
title: "20240514_HCAI_ED_Asthma_2015"
author: "Valentina Arputhasamy"
date: '2024-05-14'
output: html_document
---

This version of the code that processes the raw HCAI for use in the RR/SMR calculations was specifically created to compare our results with the tracking team. It only uses the raw HCAI data from 2015.  

```{r setup, include=FALSE}

library(writexl)
library(readxl)
library(knitr)
library(plyr)
library(reshape2)
library(ggplot2)
library(openair)
library(tidyverse)

## Diagnostic Codes: https://data.chhs.ca.gov/dataset/hospital-emergency-department-diagnosis-procedure-and-external-cause-codes

## November 8, 2023: 

# This code was developed to prepare the HCAI ED Asthma data for analysis. We will stratify by age (0-19; 20 - 44; 45 - 64; 65+) and sex (M, F) and merge 2016 with 2017 to sum total Asthma ED visits by zip. ED visits in the ED file are treat and release only and do not include the ED visits that get hospitalized. In order to include the ED to hospitalized patients, we will need to extract PDD patients using the "srcroute_ns" variable which we don't have. Therefore, this version of the code DOES NOT include PDD records. We also do not have all variables to clean the data.
```

## Data Prep 

The goal of this code is to identify all primary Asthma ED visits diagnosis from 2015 - 2017 HCAI ED data. We load in the 2015 and subset the to only include Asthma diagnoses. We then reformat the data. 

```{r setup, include=FALSE}

## Load Data and Subset to Health Outcome of Interest ##

# 2015_ED

file.2015 <- "E:/Data/Eisenman_HeatMap/Data Received July 2021/EDD/dm_ED2015.csv"

ED2015ALL <- read.csv(file.2015, stringsAsFactors=FALSE, header=T) 

ED2015ALL$date <- as.POSIXct(strptime(ED2015ALL$serv_dt, format = "%m/%d/%Y", 
                                      tz = "GMT"))

# Split the years since ICD9 was used prior to 10/1/15 and ICD10 on and after 10/1/15
ED2015_1 <- subset((selectByDate(ED2015ALL, month = c(1:9))), 
                   grepl("^493", dx_prin))
ED2015_2 <- subset((selectByDate(ED2015ALL, month = c(10:12))), 
                   grepl("^J45", dx_prin))
ED2015_Asthma <- rbind(ED2015_1, ED2015_2)
ED2015_Asthma$Year <- 2015

ED2015_Asthma <- ED2015_Asthma %>% filter(patzip >= 90000)
```

Do some recoding.

```{r setup, include=FALSE}

# Recode Age Groups

ED2015_Asthma$agyrserv <- as.numeric(ED2015_Asthma$agyrserv, 
                                     ED2015_Asthma$patzip)

ED2015_Asthma$age <- ifelse(ED2015_Asthma$agyrserv < 20, '00_19',
                            ifelse(ED2015_Asthma$agyrserv >= 20 & 
                                     ED2015_Asthma$agyrserv < 45, '20_44',
                            ifelse(ED2015_Asthma$agyrserv >= 45 & 
                                     ED2015_Asthma$agyrserv < 65, '45_64',
                            ifelse(ED2015_Asthma$agyrserv >= 65, '65+',
                            'UK'))))

ED2015_Asthma$tcases <- 1

ED2015_Asthma <- subset(ED2015_Asthma, patzip >= 90000)

mED2015_Asthma <- setNames(aggregate(ED2015_Asthma$tcases, 
                                     list(ED2015_Asthma$Year, 
                                          ED2015_Asthma$patzip, 
                                          ED2015_Asthma$age, 
                                          ED2015_Asthma$sex), 
                                     FUN=sum), 
                           c('year', 'zip', 'age', 'sex', 'tcases' ))

## get unsuppressed version of data for tracking


length(unique(mED2015_Asthma$zip)) * 4 * 2 * 3

unique_zip <- unique(mED2015_Asthma$zip)
unique_age <- unique(mED2015_Asthma$age)
unique_sex <- unique(mED2015_Asthma$sex)
unique_yr <- unique(mED2015_Asthma$year)
comb_zas15 <- expand.grid(year = unique_yr, zip = unique_zip, age = unique_age,
                          sex = unique_sex)
comb_zas15 <- subset(comb_zas15, sex != 'U')
comb_ED2015_Asthma <- merge(comb_zas15, mED2015_Asthma,
                            by = c("year", "zip", "age", "sex"), all = TRUE)
comb_ED2015_Asthma[is.na(comb_ED2015_Asthma)] <- 0
comb_ED2015_Asthma$outcome <- "ED_Asthma"
summary(comb_ED2015_Asthma$tcases)

ED2015_Asthma_AgeSex_US_tracking <- comb_ED2015_Asthma


# Save as CSV files

write.csv(ED2015_Asthma_AgeSex_US_tracking, file = 'C:/Users/VArputhasamy/Documents/HCAI_ED/Unsuppressed_Output/With_Zeros/Tracking_Data/ED2015_Asthma_AgeSex_US_fortracking.csv')
```


